### 2019-08-30 paper/paper.tex
Removed citations from abstract and made parameter table consistent.

\begin{flushleft}
{\fontsize{16pt}{20pt}\selectfont%
  Expanded Optimization for Discovering\\}
{\fontsize{16pt}{20pt}\selectfont%
  Optimal Lateral Handling Bicycles}
\end{flushleft}

@@

\section*{Abstract}
%
We introduced a method of optimizing four geometric parameters of a bicycle's
design to minimize the so called Handling Quality Metric
(HQM)~\cite{Moore2016}. Here we expand that method to optimize all X geometric
and inertial parameters of the benchmark parameterimization of the linear
Whipple-Carvallo bicycle model under constraints that guarantee a physically
realizable bicycle. This improves over the prior work by expanding the search
space with many more parameters and the guarntee of realizability.

\section{Introduction}
%
Physical design features of ground vehicles can affect their lateral handling
qualities. Geometry, mass, and mass distribution of the vehicle's primary
components as well as tire characteristics are primary contributors to poor and
good handling due to their important influence on the vehicle's dynamics. In
past work, we have presented a theoretical and computational framework for
assessing the lateral task-independent handling qualities of simplified single
track vehicle designs~\cite{Hess2012,Moore2012}. In subsequent work, we showed
that minimizing our proposed handling quality metric (HQM) can produce theoretically
optimal handling designs when only four geometric parameters are explored as
the optimization variables~\cite{Moore2016}. The present work's goal is to
expand this optimization problem to all of the geometry, mass, and inertial
parameters present in the linear Whipple-Carvallo bicycle model~\cite{Meijaard2007}.
This broadens the search space considerably but we constrain it so that only
realizable optimal bicycle designs are discovered. To do so, we formulate a
constrained optimization problem and use derivative-free optimization to
discover optimal, yet realizable, bicycle designs.

@@

parameterization''. This parameterization differs from the benchmark
parameterization in three ways.  Firstly, the person and rear frame are treated
as separate rigid bodies each with their on inertial parameters. Secondly, we
express the inertial parameters of each rigid body in terms of central
principal radii of gyration to decouple the mass from the inertia terms.
Lastly, we introduce two simple dimensional parameters that define the
geometric extents of the person which are used to constrain the location of the
person's body. Table~\ref{tab:parameters} provides the parameters and the
reference values which are derived from the measurements of a Batavus Browser
Bicycle and the rider Jason presented in~\cite{Moore2012}. This
parameterization can be transformed into the benchmark parameterization
readily, but not vice versa.

@@

  from the measurements in~\cite{Moore2012}.}

@@

@@

    $v$ &  3.0 & \si{\meter\per\second} & Forward speed \\

@@

    $m_R$ &  3.11 & \si{\kilogram} & mass \\
    $r_R$ &  0.34096 & \si{\meter} & Rear wheel radius \\
    $x_R$ & 0 & \si{\meter} & Rear wheel mass center \\
    $y_R$ & 0 & \si{\meter} & Rear wheel mass center \\
    $z_R$ & -0.34096 & \si{\meter} & Rear wheel mass center \\
    $k_{Raa}$ &  0.17050 & \si{\meter} & Rear wheel central principal radii of gyration \\
    $k_{Rbb}$ &  0.17050 & \si{\meter} & Rear wheel central principal radii of gyration  \\
    $k_{Ryy}$ &  0.22136 & \si{\meter} & Rear wheel central principal radii of gyration \\

@@

@@

    $x_F$ &  1.1210 & \si{\meter} & Mass center \\
    $y_F$ & 0.0 & \si{\meter} & Mass center \\
    $z_F$ & -0.34353 & \si{\meter} & Mass center \\
    $k_{Faa}$ &  0.20917 & \si{\meter} & Central principal radius of gyration \\
    $k_{Fbb}$ &  0.20917 & \si{\meter} & Central principal radius of gyration \\
    $k_{Fyy}$ &  0.27179 & \si{\meter} & Central principal radius of gyration \\

@@

    $l_P$ &  1.7280 & \si{\meter} & Body length \\
    $w_P$ &  0.48300 & \si{\meter} & Body width \\
    $m_P$ &  83.500 & \si{\kilogram} & Mass \\
    $x_P$ &  0.31577 & \si{\meter} & Mass center \\
    $y_P$ & 0.0 & \si{\meter} & Mass center \\
    $z_P$ & -1.0990 & \si{\meter} & Mass center \\
    $k_{Paa}$ &  0.36797 & \si{\meter} & Central principal radius of gyration \\
    $k_{Pbb}$ &  0.15276 & \si{\meter} & Central principal radius of gyration \\
    $k_{Pyy}$ &  0.36717 & \si{\meter} & Central principal radius of gyration \\
    $\alpha_P$ & 0.18618 & \si{\radian} & Principal axis angle \\

@@

    $x_H$ & 0.86695 & \si{\meter} & Mass center \\
    $y_H$ & 0.0 & \si{\meter} & Mass center \\
    $z_H$ & -0.74824 & \si{\meter} & Mass center \\
    $k_{Haa}$ & 0.29556 & \si{\meter} & Central principal radius of gyration \\
    $k_{Hbb}$ & 0.14493 & \si{\meter} & Central principal radius of gyration \\
    $k_{Hyy}$ & 0.27630 & \si{\meter} & Central principal radius of gyration \\
    $\alpha_H$ & 0.36995 & \si{\radian} & Principal axis angle \\

@@

    $x_D$ &  0.27595 & \si{\meter} & Mass center \\
    $y_D$ & 0 & \si{\meter} & Mass center \\
    $z_D$ & -0.53784 & \si{\meter} & Mass center \\
    $k_{Daa}$ &  0.28587 & \si{\meter} & Central principal radius of gyration \\
    $k_{Dbb}$ &  0.22079 & \si{\meter} & Central principal radius of gyration \\
    $k_{Dyy}$ &  0.36539 & \si{\meter} & Central principal radius of gyration \\
    $\alpha_D$ &  1.1722 & \si{\radian} & Principal axis angle \\

@@

\section{Results}
%
We discover four bicycles for four different design speeds (3, 5, 7, and 9 m/s)
that have an optimally low HQM, see Table~\ref{tab:hqm} and satisify all
constraints and parameter boundaries. We belive these bicycles to be physically
realizable with minor differences. The most striking feature is that all of the
bicycle are larger that the reference bicycle in some way.  The 3 m/s bicycle
places both the rider's mass center and the rear frame's mass center close to 3
meters above the ground plane. The 5 and 7 m/s bicycles have wheelbase values
that are simlar to the reference bicycle, but 3 and 9 m/s have very large wheel
bases. The 9 m/s bicycle is very large overall. Only the 5 m/s bicycle is of a
scale close to a typical bicycle. It is noteable that the 9 m/s bicycle has
signficant negative trail.
%
\begin{figure}
  \centering
  \includegraphics[width=6in]{../figures/optimal-geometries.png}
  \caption{Depcitions of the bicycle geometry and geometric representations of
    the inertial quanties for the reference bicycle and four optimal solutions
    at 3, 5, 7, and 9 m/s. Five rigid bodies are shown for each bicycle: front
    wheel (oragne), rear wheel (purple), rear frame (blue), front frame
    (green), and person (red). The solid black lines represent the essential
    bicycle geometry. The dotted black line represents the steer axis. The
    solid colored curves represent the contours of solid ellipsoids with
    equivalent inertia as the principal inertia of the associated rigid body.
    The dotted colored lines represent the extents of the centerial radii of
    gyration of each rigid body.}
\end{figure}
%
\begin{table}
  \caption{Peak HQM values for the reference bicycle and the optimal bicycles
    at each speed.}
  \label{tab:hqm}
  \centering
  \begin{tabular}{llll}
    \toprule
    Speed [m/s] & Reference Peak HQM & Optimal Peak HQM & Percent Improvement \\
    \midrule
    3 & 13.0753 & 2.0118 & 85\% \\
    5 & 4.5213  & 0.0115 & 100\% \\
    7 & 3.0434  & 0.0220 &  99\% \\
    9 & 2.3377  & 0.8386 &  64\% \\
    \bottomrule
  \end{tabular}
\end{table}

>>>>

\begin{flushleft}
{\fontsize{16pt}{20pt}\selectfont%
  Expanded Optimization for Discovering Optimal Lateral Handling Bicycles}
\end{flushleft}

@@

\section*{Abstract}
%
Previously, we introduced a method of optimizing four primary geometric
parameters of a bicycle's design to minimize its Handling Quality Metric. Here
we expand that method to optimize 23 geometric and inertial parameters in the
benchmark parameterimization of the linear Whipple-Carvallo bicycle model. To
ensure physically realizable bicycle designs we include 7 equality constraints,
21 inequality constraints, and maximal and minamal bounds on each free
opmization parameter. This improves over the prior work by expanding the search
space with many more parameters and the guarntee of realizability. We close by
presenting five bicycle designs that have optimal later handling qualities. The
bicycles are not generally self-stable and have some ununusal characteristics.

\section{Introduction}
%
Physical design features of ground vehicles can affect their lateral handling
qualities. Geometry, mass, and mass distribution of the vehicle's components as
well as tire characteristics are primary contributors to poor and good handling
due to their important influence on the vehicle's dynamics. In past work, we
have presented a theoretical and computational framework for assessing the
lateral task-independent handling qualities of simplified single track vehicle
designs~\cite{Hess2012,Moore2012}. In subsequent work, we showed that
minimizing our proposed handling quality metric (HQM) can produce theoretically
optimal handling designs when only four geometric parameters are explored as
the optimization variables~\cite{Moore2016}. The present work's goal is to
expand this optimization problem to all of the geometry, mass, and inertial
parameters present in the linear Whipple-Carvallo bicycle
model~\cite{Meijaard2007}.  This broadens the search space considerably but we
constrain it so that only realizable optimal bicycle designs are discovered. To
do so, we formulate a constrained optimization problem and use derivative-free
optimization to discover optimal, yet realizable, bicycle designs.

@@

parameterization'' in~\cite{Meijaard2007}. This parameterization differs from
the benchmark parameterization in three ways. Firstly, the person and rear
frame are treated as separate rigid bodies each with their on inertial
parameters. Secondly, we express the inertial parameters of each rigid body in
terms of central principal radii of gyration to decouple the mass from the
inertia terms.  Lastly, we introduce two simple dimensional parameters that
define the geometric extents of the person which are used to constrain the
location of the person's body. Table~\ref{tab:parameters} provides the
parameters and the reference values which are derived from the measurements of
a Batavus Browser Bicycle and the rider ``Jason'' presented
in~\cite{Moore2012}. This parameterization can be transformed into the
benchmark parameterization readily, but not vice versa.

@@

    from the measurements in~\cite{Moore2012} of the Batavus Browser bicycle
    and rider ``Jason''.}

@@

    % 5

@@

    $v$       &  3.0, 5.0, 7.0, 9.0 & \si{\meter\per\second} & Forward speed \\

@@

    % 8
    $m_R$     & 3.11 & \si{\kilogram}  & Mass \\
    $r_R$     & 0.34096 & \si{\meter}  & Radius \\
    $x_R$     & 0.0 & \si{\meter}      & $X$ mass center coordinate \\
    $y_R$     & 0.0 & \si{\meter}      & $Y$ mass center coordinate \\
    $z_R$     & -0.34096 & \si{\meter} & $Z$ mass center coordinate \\
    $k_{Raa}$ & 0.17050 & \si{\meter}  & Central principal radii of gyration about $A_R$ \\
    $k_{Rbb}$ & 0.17050 & \si{\meter}  & Central principal radii of gyration about $B_R$ \\
    $k_{Ryy}$ & 0.22136 & \si{\meter}  & Central principal radii of gyration about $Y$ \\

@@

    % 8

@@

    $x_F$     & 1.1210 & \si{\meter} & $X$ mass center coordinate \\
    $y_F$     & 0.0 & \si{\meter} & $Y$ mass center coordinate \\
    $z_F$     & -0.34353 & \si{\meter} & $Z$ mass center coordinate \\
    $k_{Faa}$ &  0.20917 & \si{\meter} & Central principal radii of gyration about $A_F$ \\
    $k_{Fbb}$ &  0.20917 & \si{\meter} & Central principal radii of gyration about $B_F$ \\
    $k_{Fyy}$ &  0.27179 & \si{\meter} & Central principal radii of gyration about $Y$ \\

@@

    % 10
    $l_P$      &  1.728   & \si{\meter} & Body length \\
    $w_P$      &  0.483   & \si{\meter} & Body width \\
    $m_P$      &  83.5    & \si{\kilogram} & Mass \\
    $x_P$      &  0.31577 & \si{\meter} & $X$ mass center coordinate \\
    $y_P$      & 0.0      & \si{\meter} & $Y$ mass center coordinate \\
    $z_P$      & -1.0990  & \si{\meter} & $Z$ mass center coordinate \\
    $k_{Paa}$  &  0.36797 & \si{\meter} & Central principal radii of gyration about $A_P$ \\
    $k_{Pbb}$  &  0.15276 & \si{\meter} & Central principal radii of gyration about $B_P$ \\
    $k_{Pyy}$  &  0.36717 & \si{\meter} & Central principal radii of gyration about $Y$ \\
    $\alpha_P$ & 0.18618 & \si{\radian} & Angle about $Y$ between $A_P$ and $X$ \\

@@

    $x_H$ & 0.86695 & \si{\meter}       & $X$ mass center coordinate \\
    $y_H$ & 0.0 & \si{\meter}           & $Y$ mass center coordinate \\
    $z_H$ & -0.74824 & \si{\meter}      & $Z$ mass center coordinate \\
    $k_{Haa}$ & 0.29556 & \si{\meter}   & Central principal radii of gyration about $A_H$ \\
    $k_{Hbb}$ & 0.14493 & \si{\meter}   & Central principal radii of gyration about $B_H$ \\
    $k_{Hyy}$ & 0.27630 & \si{\meter}   & Central principal radii of gyration about $Y$ \\
    $\alpha_H$ & 0.36995 & \si{\radian} & Angle about $Y$ between $A_H$ and $X$ \\

@@

    $x_D$ &  0.27595 & \si{\meter}      & $X$ mass center coordinate \\
    $y_D$ & 0 & \si{\meter}             & $Y$ mass center coordinate \\
    $z_D$ & -0.53784 & \si{\meter}      & $Z$ mass center coordinate \\
    $k_{Daa}$ &  0.28587 & \si{\meter}  & Central principal radii of gyration about $A_D$ \\
    $k_{Dbb}$ &  0.22079 & \si{\meter}  & Central principal radii of gyration about $B_D$ \\
    $k_{Dyy}$ &  0.36539 & \si{\meter}  & Central principal radii of gyration about $Y$ \\
    $\alpha_D$ &  1.1722 & \si{\radian} & Angle about $Y$ between $A_D$ and $X$ \\

@@

\section{Results}
%
We discover four bicycles for four different design speeds (3, 5, 7, and 9 m/s)
that have an optimally low HQM, see Table~\ref{tab:hqm} and satisify all
constraints and parameter boundaries. We belive these bicycles to be physically
realizable with minor differences. The most striking feature is that all of the
bicycle are larger that the reference bicycle in some way.  The 3 m/s bicycle
places both the rider's mass center and the rear frame's mass center close to 3
meters above the ground plane. The 5 and 7 m/s bicycles have wheelbase values
that are simlar to the reference bicycle, but 3 and 9 m/s have very large wheel
bases. The 9 m/s bicycle is very large overall. Only the 5 m/s bicycle is of a
scale close to a typical bicycle. It is noteable that the 9 m/s bicycle has
signficant negative trail.

3
- Rider is over 3 m in the air and leaned forward.
- 3 m wheel base with 0.7 meter positive trail
- wheels are over double the diameter of browser wheels
- minimum front frame mass

5 m/s
- rider is laid back a 0.5 meter off ground
- < 1 m wheelbase
- essentially zero trail (5 mm negative)
- minimum front frame mass

7 m/s
- miminal rotatoinal inertia about the steer axis
- tall slender fork/handlebar
- no trail (1 mm postive)
- half sized wheels (~12 inches)
- normal wheelbase (1.157)
- rider tipped forward and low (0.6 m)
- vertical steer axis
- 3 m high frame center of mass

9
- very large 5.6 m wheelbase
%
\begin{figure}
  \centering
  \includegraphics[width=6in]{../figures/optimal-geometries.png}
  \caption{Depcitions of the bicycle geometry and geometric representations of
    the inertial quanties for the reference bicycle and four optimal solutions
    at 3, 5, 7, and 9 m/s. Five rigid bodies are shown for each bicycle: front
    wheel (oragne), rear wheel (purple), rear frame (blue), front frame
    (green), and person (red). The solid black lines represent the essential
    bicycle geometry. The dotted black line represents the steer axis. The
    solid colored curves represent the contours of solid ellipsoids with
    equivalent inertia as the principal inertia of the associated rigid body.
    The dotted colored lines represent the extents of the centerial radii of
    gyration of each rigid body.}
\end{figure}
%
\begin{table}
  \caption{Peak HQM values for the reference bicycle and the optimal bicycles
    at each speed.}
  \label{tab:hqm}
  \centering
  \begin{tabular}{llll}
    \toprule
    Speed [m/s] & Reference Peak HQM & Optimal Peak HQM & Percent Improvement \\
    \midrule
    3 & 13.0753 & 2.0118 & 85\% \\
    5 & 4.5213  & 0.0115 & 100\% \\
    7 & 3.0434  & 0.0220 &  99\% \\
    9 & 2.3377  & 0.8386 &  64\% \\
    \bottomrule
  \end{tabular}
\end{table}
%
\begin{figure}
  \centering
  \includegraphics[width=6in]{../figures/optimal-eigenvalues.png}
  \caption{}
\end{figure}
%
\begin{table}
  \caption{Optimal principal parameter values for each design speed.}
  \label{tab:optimal-values}
  \centering
  \begin{tabular}{lSSSS}
    \toprule
    Parameter & 3~\si{\meter\per\second} & 5~\si{\meter\per\second} & 7~\si{\meter\per\second} & 9~\si{\meter\per\second} \\
    \midrule
    $c$       & 0.688  & -0.005 & 0.001  & -0.484 \\
    $w$       & 2.866  & 0.847  & 1.157  & 5.557  \\
    $\lambda$ & -0.213 & 0.271  & -0.028 & 0.239  \\
    \midrule
    $m_D$ & 3.22   &  11.53 & 10.32  & 7.02 \\
    $x_D$ & 0.958  &  0.287 &  0.427 & 1.020 \\
    $z_D$ & -2.605 & -2.719 & -2.976 & -4.065 \\
    $\alpha_D$ & 0.663  & 0.915  & 1.123  & 1.077 \\
    % Should aa be larger than bb?
    $k_{Daa}$  & 1.449  & 1.218  &  1.324 & 0.861 \\
    $k_{Dbb}$  & 0.724  & 1.387  &  1.290 & 2.014 \\
    $k_{Dyy}$  & 1.471  & 1.559  &  1.555 & 2.031 \\
    \midrule
    $m_H$  & 0.25  & 0.25 &    0.49 &   3.54 \\
    $x_H$     & 2.356   & 0.532 &    1.145 &    5.615 \\
    $z_H$     & -1.298  & -0.781 &   -1.340 &   -1.776 \\
    $\alpha_H$ & 0.572  & -1.265 & 1.571  & 0.238 \\
    $k_{Haa}$ & 0.186  & 0.0491 &   0.000 &       2.846 \\
    $k_{Hbb}$ & 1.259   & 1.145 &     2.006 &    0.168 \\
    $k_{Hyy}$ & 0.636  & 0.383 &    0.818 &   0.669 \\
    \midrule
    $m_F$     & 1.62  & 4.40  & 2.64  & 5.51  \\
    $r_F$     & 0.710 & 0.252 & 0.127 & 2.063 \\
    $k_{Faa}$ & 0.355 & 0.126 & 0.064 & 1.031 \\
    $k_{Fyy}$ & 0.710 & 0.252 & 0.127 & 2.063 \\
    \midrule
    $x_P$      & 0.765  & 0.276  &  0.526 & 1.079 \\
    $z_P$      & -3.194 & -0.453 & -0.586 &-2.662 \\
    $\alpha_P$ & -0.661 & 1.150  & -0.836 & 1.558 \\
    \midrule
    $m_R$   & 12.63  & 8.36  &  6.31 & 2.25 \\
    $r_R$   & 0.958  & 0.506 & 0.146 & 3.027 \\
    $k_Raa$ & 0.479  & 0.253 & 0.073 & 1.514 \\
    $k_Ryy$ & 0.958  & 0.506 & 0.146 & 3.027 \\
    \bottomrule
  \end{tabular}
\end{table}
